---
title: "Data 607 Assignment 9 – Web APIs"
author: "Leslie Tavarez"
date: "2024-11-01"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
In this assignment, I will use The New York Times APIs, which provide a variety of data. First, I will sign up for an API key. Then, I’ll choose one API to work with and create an interface in R to fetch and read the JSON data from it. Finally, I will convert this data into an R DataFrame, making it easier to analyze and manipulate. This will help me learn how to work with web APIs and handle data in R.


##### API Connection 
```{r}
library(httr)
library(jsonlite)
library(dplyr)


# API Key
apikey <- "IsXIAthPjA6e9jFJJevYb5xrlSfecGDf"

# Construct the URL with the API key as a query parameter
theURL <- paste0("https://api.nytimes.com/svc/topstories/v2/world.json?api-key=", apikey)

# Make the GET request
stories <- GET(theURL)

# Check the status code
print(stories$status_code)

```
##### Converting API into DataFrame
```{r}
summary(stories)

#Parse data
parsed_data <- content(stories, "text", encoding = "UTF-8")
stories_df <- fromJSON(parsed_data)

#Extract relevant parts of the data
  articles <- stories_df$results
  
#Convert to DataFrame
  articles_df <- as.data.frame(articles, stringsAsFactors = FALSE)
  
# Rename the byline column to author
articles_df <- articles_df %>%
  rename(author = byline)

# Remove "By " from the author column
articles_df$author <- gsub("^By ", "", articles_df$author)

# Select and print the title and author columns
selected_columns <- articles_df[, c("title", "author")]
print(selected_columns)
```

In this assignment, I faced challenges when attempting to create a DataFrame before properly parsing the JSON data. Parsing is essential when working with JSON because it converts the structured text into a format that R can easily manipulate. Through this process, I learned the importance of extracting relevant information from the API response and transforming it into a usable DataFrame. 